---
title: Example II
description: ''
date: 2025-06-04
weight: 6
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo=FALSE, include=FALSE}
source("R/site_globals.R")
source("scripts/geomUtils.R")
source("scripts/config.R")
library(leaflet)
library(rmarkdown)
library(dplyr)
library(ggplot2)
```

## Historical habitat

This map shows the accumulated search effort for Crassula connata over its historical habitat covering 6 grid cells
in Bellhouse Park:

```{r, error=TRUE}
cracon_agm <- read.csv("Analysis_outputs/Intermediate/Crassula connata_accepted_grouped_merged.csv")
cracon_historical <- cracon_agm %>% dplyr::filter(assigned_community == "CC1")
cracon_sf = assign_cell_geometry_sf(cracon_historical, galgrid)

pal <- colorNumeric(palette = "viridis", domain = range(c(0, cracon_sf$search_effort), na.rm = TRUE))
m <- leaflet(data = cracon_sf) %>%
  # Add a Tiles layer to the map
  addProviderTiles("Esri.WorldImagery") %>%
  # Add the grid layer to the map
  addPolygons(fillColor = ~pal(search_effort), fillOpacity = 0.8, 
              color = "#BDBDC3", weight = 1) %>%
  # Add a legend
  addLegend(pal = pal, values = c(0, max(cracon_sf$search_effort, na.rm = TRUE)),
            opacity = 0.8, title = "Accumulated Search Effort in ks")

# Print the map
m

```

The table that this plot is generated from is in the AGM (accepted, grouped, merged) format, output by the 
data pipeline which aggregates search trace data, grids it into cells and intersects it with habitat polygons. Here is the data covering
the 6 Bellhouse Park cells:

```{r, message=FALSE, echo=FALSE}
cracon_historical_trunc <- cracon_historical %>% trim_effortId
paged_table(cracon_historical_trunc)
```

This data format is explained in the [previous example](../example-i)

## Potential habitat 


This map shows the accumulated search effort for Crassula connata over both historical and potential habitat:

```{r}
cracon_agm <- read.csv("Analysis_outputs/Intermediate/Crassula connata_accepted_grouped_merged.csv")
cracon_sf = assign_cell_geometry_sf(cracon_agm, galgrid)

pal <- colorNumeric(palette = "viridis", domain = range(c(0, cracon_sf$search_effort), na.rm = TRUE))
m <- leaflet(data = cracon_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  # Add the grid layer to the map
  addPolygons(fillColor = ~pal(search_effort), fillOpacity = 0.8, 
              color = "#BDBDC3", weight = 1) %>%
  # Add a legend
  addLegend(pal = pal, values = c(0, max(cracon_sf$search_effort, na.rm = TRUE)),
            opacity = 0.8, title = "Accumulated Search Effort in ks")

# Print the map
m

```

See the treatment of [historical habitat](../../extirpation_historical/example-i) for explanation of the data format underlying this plot.
