---
title: Rarity informed searches
description: 
math: true
weight: 1
---

### Crassula

```{r, warning=FALSE}
cracon_accepted_sf <- st_read("Analysis_outputs/Crassula connata_Solow_low.shp", quiet=TRUE)
# Filter for region
# Convert from mean likelihood of presence to mean likelihood of extirpation
cracon_accepted_sf$mean_ep <- 1 - cracon_accepted_sf$mean

pal <- colorNumeric(palette = "viridis", domain = range(c(0, cracon_accepted_sf$mean_ep), na.rm = TRUE))
m <- leaflet(data = cracon_accepted_sf) %>%
  # Add a Tiles layer to the map
  addProviderTiles("Esri.WorldImagery") %>%
  # Add the grid layer to the map
  addPolygons(fillColor = ~pal(mean_ep), fillOpacity = 0.8, 
              color = "#BDBDC3", weight = 1) %>%
  # Add a legend
  addLegend(pal = pal, values = c(0, max(cracon_accepted_sf$mean_ep, na.rm = TRUE)),
            opacity = 0.8, title = "Mean likelihood of extirpation")

# Print the map
m

```

### Primula

```{r, warning=FALSE}
prim_accepted_sf <- st_read("Analysis_outputs/Primula pauciflora_Solow_low.shp", quiet=TRUE)
# Filter for region
# Convert from mean likelihood of presence to mean likelihood of extirpation
prim_accepted_sf$mean_ep <- 1 - prim_accepted_sf$mean

pal <- colorNumeric(palette = "viridis", domain = range(c(0, prim_accepted_sf$mean_ep), na.rm = TRUE))
m <- leaflet(data = prim_accepted_sf) %>%
  # Add a Tiles layer to the map
  addProviderTiles("Esri.WorldImagery") %>%
  # Add the grid layer to the map
  addPolygons(fillColor = ~pal(mean_ep), fillOpacity = 0.8, 
              color = "#BDBDC3", weight = 1) %>%
  # Add a legend
  addLegend(pal = pal, values = c(0, max(prim_accepted_sf$mean_ep, na.rm = TRUE)),
            opacity = 0.8, title = "Mean likelihood of extirpation")

# Print the map
m

```

| Species                   | Rarity / Habitat Notes |
|---------------------------|-------------------------|
| *Crassula connata*        | Very low occupancy. Typically found in localised patches within a habitat patch; absent from most islands; widely dispersed across the larger islands of Vancouver and San Juan. |
| *Meconella oregana*       | Very low occupancy. Absent from most islands; moderately distributed on the southern tip of Vancouver Island but only known from single localities on other islands where it is highly local within a patch when found. |
| *Perideridia montana*     | Low occupancy. Widely distributed on the southern tip of Vancouver Island, but idiosyncratic distribution on smaller islands; often locally abundant and widely distributed within habitat patches when found; sometimes restricted to a narrower microhabitat (swales) in drier sites. |
| *Trifolium dichotomum*    | Low occupancy. Found on a few islands where records are highly clustered; locally abundant to common in patches where found, not widely distributed within patches when found. |
| *Aquilegia formosa*       | Low occupancy. Mostly found on larger islands where populations are spatially clustered, though widely distributed; preference as deer forage may mask distribution. |
| *Castilleja attenuata*    | Low occupancy. Sparsely distributed on smaller islands but widely distributed on the southern tip of Vancouver Island. |
| *Plagiobothrys tenellus*  | Low occupancy. Sparsely distributed on larger islands; often locally abundant and widely distributed within habitat patches when found. |
| *Platanthera unalascensis*| Moderate occupancy. Sparsely distributed on larger islands but widely distributed on the southern tip of Vancouver Island. |
| *Lepidium virginicum*     | Moderate occupancy. Fairly broadly distributed across islands and habitat patches within islands, though infrequently occurring. |
| *Primula pauciflora*      | Moderate occupancy. Found on a few widely scattered islands; localities highly clustered within islands, where it is often locally common and widely distributed within patches when found. |

{{% imgproc Crassula_connata Fill "2400x600" %}}
Crassula connata (Image credit: Daniel
Hetzel)
{{% /imgproc %}}

XXXXX
