---
title: Search effort
description: Incorporating search effort
weight: 5
sidebar:
  open: true
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo=FALSE, include=FALSE}
source("R/site_globals.R")
source("scripts/plotTraces.R")
```

Search effort is abstracted from two kinds of observation data: 

* GPS traces collected from handheld GPS receivers carried by the observers
* Traces inferred from iNaturalist observation data by interpolating tracks imputing travel at constant speed between successive observations made by each observer participating in a search event

The following map shows blue markers for each iNaturalist observation made in Bellhouse Park by observers participating in this study, interpolated with white circles connecting them into traces:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
bellhouseObsMap
```

The code performing the interpolation is in [synthesizeSearchEffort.R](`r site_rel("/scripts/synthesizeSearchEffort.R")`)
which loops over the tabular data supplied in [Search_Effort_Summary.csv](`r site_rel("/Analysis_inputs/Search_Effort/Search_Effort_Summary.csv")`) as follows:
```
for (parsedEffortId in effortWithObsIds) {
...
    for (trackObserver in trackObservers) {
    ...
          newTrace_sf <- trackToTrace(parsedEffortId, obsFilteredTrack_sf, anEffort, obsTraceEvery)
}
```

Gathering these interpolated traces together with GPS tracks produces the following combined trace map for the region:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
allTracesMap
```

The effort encoded in these traces is then intersected with the
30x30 m habitat cells chosen for the study, producing a grid of search effort density:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
bellhouseSearchMap
```

Search effort is represented as per-observer search time (in kiloseconds) per gridded cell. 
Search effort applies only to valid potential or historical habitat for the target species. 
Where habitat mapping only overlaps partially with the grid, the imputed search effort
is scaled up proportionally.
