---
title: Search effort
description: Incorporating search effort
weight: 5
sidebar:
open: true
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo=FALSE, include=FALSE}
source("R/site_globals.R")
source("scripts/plotTraces.R")
```

Search effort is assembled from two kinds of observation data - 

* GPS traces collected from handheld GPS receivers carried by the observers
* Traces inferred from iNaturalist observation data by interpolating tracks imputing travel at constant speed between successive
observations made by each observer participating in a search effort event.

The following map shows blue markers for each iNaturalist observation made in Bellhouse Park by observers participating
in this study, together with interpolated traces as white circles connecting them into traces:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
bellhouseObsMap
```

The code performing the interpolation is in [geomUtil.R](`r site_rel("/scripts/synthesizeSearchEffort.R")`) which loops over the tabular data
supplied in [Search_Effort_Summary.csv](`r site_rel("/Analysis_inputs/Search_Effort/Search_Effort_Summary.csv")`) as follows:
```
for (parsedEffortId in effortWithObsIds) {
...
    for (trackObserver in trackObservers) {
    ...
          newTrace_sf <- trackToTrace(parsedEffortId, obsFilteredTrack_sf, anEffort, obsTraceEvery)
}
```

Collecting together all such interpolated traces together with raw traces collected by observers with
dedicated GPS receivers produces the following combined trace map for the region:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
allTracesMap
```

Finally effort from all observers encoded in such traces is gridded with respect to the standard
30x30 m cells chosen for the study, leading to the following search effort density:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
bellhouseSearchMap
```


Search effort is represented as a per-observer search time per gridded cell
represented in ksec. Search effort is applied only to valid potential
or historical habitat for the target taxon. Where the habitat model indicates
that only part of a grid cell is valid habitat, the imputed search effort
is scaled up proportionally.
