---
title: Example II
description: ''
date: 2025-06-04
weight: 6
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo=FALSE, include=FALSE}
source("R/site_globals.R")
source("scripts/geomUtils.R")
source("scripts/config.R")
library(leaflet)
library(rmarkdown)
library(dplyr)
library(ggplot2)
```

## Historical habitat

This map shows the accumulated search effort for *Crassula connata* over its historical habitat covering 6 grid cells in Bellhouse Park:

```{r, echo=FALSE}
cracon_agm <- read.csv("Analysis_outputs/Intermediate/Crassula connata_accepted_grouped_merged.csv")
cracon_historical <- cracon_agm %>% dplyr::filter(assigned_community == "CC1")
cracon_sf = assign_cell_geometry_sf(cracon_historical, galgrid)

cracon_sf$scaled_se = cracon_sf$search_effort / cracon_sf$area_prop

eps <- 1e-6

# Scale legend palette for legibility

f_transform <- function(x, eps = 1e-6) {
  log(1 + log(x + eps))
}

f_inverse <- function(y) {
  exp(exp(y) -1)
}

min_val <- min(c(cracon_sf$search_effort, cracon_sf$scaled_se), na.rm = TRUE)
max_val <- max(c(cracon_sf$search_effort, cracon_sf$scaled_se), na.rm = TRUE)

domain_trans <- f_transform(c(min_val, max_val))

pal <- colorNumeric(
  palette = "viridis",
  domain = domain_trans
)

m <- leaflet(data = cracon_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(
    fillColor = ~pal(f_transform(search_effort)),
    fillOpacity = 0.8,
    color = "#BDBDC3",
    weight = 1
  ) %>%
  addLegend(
    pal = pal,
    values = domain_trans,
    labFormat = leaflet::labelFormat(
      transform = function(y) f_inverse(y),
      digits = 2
    ),
    opacity = 0.8,
    title = "Accumulated Search Effort in ks"
  )

m

```

```{r, eval=FALSE}
cracon_agm <- read.csv("Analysis_outputs/Intermediate/Crassula connata_accepted_grouped_merged.csv")
cracon_historical <- cracon_agm %>% dplyr::filter(assigned_community == "CC1")
cracon_sf = assign_cell_geometry_sf(cracon_historical, galgrid)

cracon_sf$scaled_se = cracon_sf$search_effort / cracon_sf$area_prop

eps <- 1e-6

# Scale legend palette for legibility

f_transform <- function(x, eps = 1e-6) {
  log(1 + log(x + eps))
}

f_inverse <- function(y) {
  exp(exp(y) -1)
}

min_val <- min(c(cracon_sf$search_effort, cracon_sf$scaled_se), na.rm = TRUE)
max_val <- max(c(cracon_sf$search_effort, cracon_sf$scaled_se), na.rm = TRUE)

domain_trans <- f_transform(c(min_val, max_val))

pal <- colorNumeric(
  palette = "viridis",
  domain = domain_trans
)

m <- leaflet(data = cracon_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(
    fillColor = ~pal(f_transform(search_effort)),
    fillOpacity = 0.8,
    color = "#BDBDC3",
    weight = 1
  ) %>%
  addLegend(
    pal = pal,
    values = domain_trans,
    labFormat = leaflet::labelFormat(
      transform = function(y) f_inverse(y),
      digits = 2
    ),
    opacity = 0.8,
    title = "Accumulated Search Effort in ks"
  )

m

```

Some cells at the fringe of the habitat patch overlap only partially with valid habitat; within these cells, search effort should not scale to the full 30 × 30 m cell area. We therefore scale search effort in proportion to the area of habitat contained within each cell.

```{r, echo=FALSE}
m <- leaflet(data = cracon_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(
    fillColor = ~pal(f_transform(scaled_se)),
    fillOpacity = 0.8,
    color = "#BDBDC3",
    weight = 1
  ) %>%
  addLegend(
    pal = pal,
    values = domain_trans,
    labFormat = leaflet::labelFormat(
      transform = function(y) f_inverse(y),
      digits = 2
    ),
    opacity = 0.8,
    title = "Accumulated Search Effort (ks)"
  )

m

```

```{r, eval=FALSE}
m <- leaflet(data = cracon_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(
    fillColor = ~pal(f_transform(scaled_se)),
    fillOpacity = 0.8,
    color = "#BDBDC3",
    weight = 1
  ) %>%
  addLegend(
    pal = pal,
    values = domain_trans,
    labFormat = leaflet::labelFormat(
      transform = function(y) f_inverse(y),
      digits = 2
    ),
    opacity = 0.8,
    title = "Accumulated Search Effort (ks)"
  )

m

```

This plot is based on an AGM table (accepted, grouped, merged) produced by our workflow that aggregates the search-trace data, grids it into cells, and intersects those cells with the habitat polygons. Shown here is the subset of data for the six Bellhouse Park cells:

```{r, message=FALSE, echo=FALSE}
cracon_historical_trunc <- cracon_historical %>% trim_effortId
paged_table(cracon_historical_trunc)
```

This data format is explained in the [previous example](../example-i)

## Potential habitat 


This map shows accumulated search effort targeting *Crassula connata* throughout both historical and potential habitat:

```{r}
cracon_agm <- read.csv("Analysis_outputs/Intermediate/Crassula connata_accepted_grouped_merged.csv")
cracon_sf = assign_cell_geometry_sf(cracon_agm, galgrid)

pal <- colorNumeric(palette = "viridis", domain = range(c(0, cracon_sf$search_effort), na.rm = TRUE))
m <- leaflet(data = cracon_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  # Add the grid layer to the map
  addPolygons(fillColor = ~pal(search_effort), fillOpacity = 0.8, 
              color = "#BDBDC3", weight = 1) %>%
  # Add a legend
  addLegend(pal = pal, values = c(0, max(cracon_sf$search_effort, na.rm = TRUE)),
            opacity = 0.8, title = "Accumulated Search Effort in ks")

# Print the map
m

```

See the vignette covering inference of extirpation in [historical habitat](../../extirpation_historical/example-i) for further explanation of the data format underlying this plot.
