---
title: Example II
description: ''
date: 2025-06-04
weight: 6
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo=FALSE, include=FALSE}
source("R/site_globals.R")
source("scripts/geomUtils.R")
source("scripts/config.R")
library(leaflet)
library(rmarkdown)
library(dplyr)
library(ggplot2)
```

## Historical habitat

This map shows the accumulated search effort for Crassula connata over its historical habitat covering 6 grid cells
in Bellhouse Park:

```{r, error=TRUE}
cracon_agm <- read.csv("Analysis_outputs/Intermediate/Crassula connata_accepted_grouped_merged.csv")
cracon_historical <- cracon_agm %>% dplyr::filter(assigned_community == "CC1")
cracon_sf = assign_cell_geometry_sf(cracon_historical, galgrid)

cracon_sf$scaled_se = cracon_sf$search_effort / cracon_sf$area_prop

eps <- 1e-6

# Scale legend palette for legibility

f_transform <- function(x, eps = 1e-6) {
  log(1 + log(x + eps))
}

f_inverse <- function(y) {
  exp(exp(y) -1)
}

min_val <- min(c(cracon_sf$search_effort, cracon_sf$scaled_se), na.rm = TRUE)
max_val <- max(c(cracon_sf$search_effort, cracon_sf$scaled_se), na.rm = TRUE)

domain_trans <- f_transform(c(min_val, max_val))

pal <- colorNumeric(
  palette = "viridis",
  domain = domain_trans
)

m <- leaflet(data = cracon_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(
    fillColor = ~pal(f_transform(search_effort)),
    fillOpacity = 0.8,
    color = "#BDBDC3",
    weight = 1
  ) %>%
  addLegend(
    pal = pal,
    values = domain_trans,
    labFormat = leaflet::labelFormat(
      transform = function(y) f_inverse(y),
      digits = 2
    ),
    opacity = 0.8,
    title = "Accumulated Search Effort in ks"
  )

m

```

The gridded cells at the fringe of the range have only a small overlap with valid habitat and so should not have their search effort referred to the full cell size of 30x30m^2. Scaling the search effort per gridded cell to take account
of this shows the search as being more complete in these areas:

```{r}
m <- leaflet(data = cracon_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(
    fillColor = ~pal(f_transform(scaled_se)),
    fillOpacity = 0.8,
    color = "#BDBDC3",
    weight = 1
  ) %>%
  addLegend(
    pal = pal,
    values = domain_trans,
    labFormat = leaflet::labelFormat(
      transform = function(y) f_inverse(y),
      digits = 2
    ),
    opacity = 0.8,
    title = "Accumulated Search Effort (ks)"
  )

m

```


The table that this plot is generated from is in the AGM (accepted, grouped, merged) format, output by the 
data pipeline which aggregates search trace data, grids it into cells and intersects it with habitat polygons. Here is the data covering
the 6 Bellhouse Park cells:

```{r, message=FALSE, echo=FALSE}
cracon_historical_trunc <- cracon_historical %>% trim_effortId
paged_table(cracon_historical_trunc)
```

This data format is explained in the [previous example](../example-i)

## Potential habitat 


This map shows the accumulated search effort for Crassula connata over both historical and potential habitat:

```{r}
cracon_agm <- read.csv("Analysis_outputs/Intermediate/Crassula connata_accepted_grouped_merged.csv")
cracon_sf = assign_cell_geometry_sf(cracon_agm, galgrid)

pal <- colorNumeric(palette = "viridis", domain = range(c(0, cracon_sf$search_effort), na.rm = TRUE))
m <- leaflet(data = cracon_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  # Add the grid layer to the map
  addPolygons(fillColor = ~pal(search_effort), fillOpacity = 0.8, 
              color = "#BDBDC3", weight = 1) %>%
  # Add a legend
  addLegend(pal = pal, values = c(0, max(cracon_sf$search_effort, na.rm = TRUE)),
            opacity = 0.8, title = "Accumulated Search Effort in ks")

# Print the map
m

```

See the treatment of [historical habitat](../../extirpation_historical/example-i) for explanation of the data format underlying this plot.
