---
title: Kernel validation
description: Insights from metapopulation theory
math: true
weight: 3
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(render_markdown = FALSE)
```

```{r, echo=FALSE, include=FALSE}
source("scripts/plotDecay.R")

```

To assess whether our dispersal kernels produce sensible predictions, we validate them for detected species; these comparisons provide a benchmark for interpreting kernel estimates in cases where no detections were made.

For detected species, negative-exponential kernels were estimated empirically by regressing the distances of species detections from the position of initial records, against the background of unsuccessful detections.

In the following graph, the black circles show the three successful detections of one of our target taxa,
_Aquilegia formosa_,  plotted with value 1, together with the background of unsuccessful detections plotted with
value 0, where the x axis shows distance of the search event from its historical habitat. The blue curve shows the
best fitted kernel value for the negative-exponential occurrence function which contributes to our prior weighting,
which rounds to a value of 0.5 in inverse km units.

```{r, message=FALSE, echo=FALSE, warning=FALSE, fig.width=10, fig.height=5}
fitplot
```

<div class="para">The fitting process was achieved with the brms package (Bürkner, 2017) using Markov Chain Monte Carlo sampling in STAN. Our regression applies the Bernoulli distribution family to the target detection via a modelling function
\(occupancy_i 〜 q .exp(-\gamma d_i) \) where \(\gamma\) is the kernel width, q scales the overall detection rate (not of interest to this analysis) and \(d_i\) is the distance from historical habitat of occupancy observation \(occupancy_i\). The following figure shows the full estimated posterior density for the kernel parameter \(\gamma\) which has a mode
at 0.32 and a 95% highest posterior density (HPD) region between \([0.07, 1.39]\):</div>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
densityplot
```
