---
title: Example II
description: ''
date: 2025-06-04
weight: 6
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo=FALSE, include=FALSE}
source("R/site_globals.R")
source("scripts/geomUtils.R")
source("scripts/config.R")
library(leaflet)
library(rmarkdown)
library(dplyr)
library(ggplot2)
```

Gridded search data are passed for computing regional statistics as the second argument to the ``analyse_accepted`` function in Analyse.R. This data format is explained in the [previous example](../example-i)

This produces posterior regional statistics for extirpation in historical habitat as follows:

```{r}
target_stats <- read.csv("Analysis_outputs/Intermediate/Crassula connata_stats.csv")
target_stats_historical <- target_stats %>% dplyr::filter(Population == "CC1")
paged_table(target_stats_historical)
```

These show the computed parameters of the posterior beta distribution for extirpation expressed in two different forms - the standard (alpha, beta) representation and (mu, var) as parameters for the central estimate for sighting probability and its dispersion. Confidence bands are placed for this at [`r target_stats_historical$Low`, `r target_stats_historical$High`].

Here is this posterior distribution graphed out:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot_site_beta(target_stats_historical, "Crassula connata")
```

