---
title: Example II
description: ''
date: 2025-06-04
weight: 6
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo=FALSE, include=FALSE}
source("R/site_globals.R")
source("scripts/geomUtils.R")
source("scripts/config.R")
library(leaflet)
library(rmarkdown)
library(dplyr)
library(ggplot2)
```

Similarly to the previous vignette, this map plots this mean per-cell posterior extirpation likelihood for _Crassula connata_ over its historical habitat covering 6 grid cells in Bellhouse Park. 

```{r, warning=FALSE}
cracon_accepted_sf <- st_read("Analysis_outputs/Crassula connata_Solow_low.shp", quiet=TRUE)
# Filter for region
cracon_historical_sf <- cracon_accepted_sf %>% dplyr::filter(assgnd_ == 44)
# Convert from mean likelihood of presence to mean likelihood of extirpation
cracon_historical_sf$mean_ep <- 1 - cracon_historical_sf$mean

pal <- colorNumeric(palette = "viridis", domain = range(c(0, cracon_historical_sf$mean_ep), na.rm = TRUE))
m <- leaflet(data = cracon_historical_sf) %>%
  # Add a Tiles layer to the map
  addProviderTiles("Esri.WorldImagery") %>%
  # Add the grid layer to the map
  addPolygons(fillColor = ~pal(mean_ep), fillOpacity = 0.8, 
              color = "#BDBDC3", weight = 1) %>%
  # Add a legend
  addLegend(pal = pal, values = c(0, max(cracon_historical_sf$mean_ep, na.rm = TRUE)),
            opacity = 0.8, title = "Mean likelihood of extirpation")

# Print the map
m

```

Gridded search data are passed for computing regional statistics as the second argument to the ``analyse_accepted`` function in Analyse.R. This data format is explained in the [previous example](../example-i)

This produces posterior regional statistics for extirpation in historical habitat as follows:

```{r}
target_stats <- read.csv("Analysis_outputs/Intermediate/Crassula connata_stats.csv")
target_stats_historical <- target_stats %>% dplyr::filter(Population == "CC1")
paged_table(target_stats_historical)
```

These show the computed parameters of the posterior beta distribution for extirpation expressed in two different forms - the standard (alpha, beta) representation and (mu, var) as parameters for the central estimate for sighting probability and its dispersion. Confidence bands are placed for this at [`r target_stats_historical$Low`, `r target_stats_historical$High`].

Here is this posterior distribution graphed out:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot_site_beta(target_stats_historical, "Crassula connata")
```

